---
title: "Course Project 1"
output: md_document
---

## Load the data into R using read.csv creating variable "Activity"

```{r load data, echo = TRUE}
Activity <- read.csv("activity.csv")
```

## First question: What is mean total number of steps taken per day?

```{r, echo = TRUE}
library(ggplot2)
totsteps <- tapply(Activity$steps, Activity$date, FUN = sum, na.rm = TRUE)
qplot(totsteps, binwidth = 1000, xlab = "Total Number of Steps Taken per Day")
mean(totsteps, na.rm = TRUE)
median(totsteps, na.rm = TRUE)
```

## Second question: What is the average daily activity pattern?

```{r, echo = TRUE}
library(ggplot2)
average_daily <- aggregate(x = list(steps = Activity$steps), by = list(interval = Activity$interval), FUN = sum, na.rm = TRUE)
ggplot(data = average_daily, aes(x = interval, y = steps)) +
  geom_line()+
  xlab("5-min Interval") + 
  ylab("Average of Steps")
```

Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r, echo = TRUE}
average_daily[which.max(average_daily$steps),]
```

## Third question: Imputing missing values

Note that there are a number of days/intervals where there are missing values (coded as NA). The presence of missing days may introduce bias into some calculations or summaries of the data.

```{r, echo = TRUE}
miss_values <- is.na(Activity$steps)
table(miss_values)
```

The strategy for filling in all of the missing values in the dataset will be the mean value of that 5-minute interval.

```{r, echo = TRUE}
# Create a new dataset that is equal to the original dataset but with the missing data filled in.
filling_values <- function(steps, interval) {
  fill <- NA
  if (!is.na(steps))
    fill <- c(steps)
  else
    fill <- (average_daily[average_daily$interval == interval, "steps"])
  return(fill)
}
fill_Activity <- Activity
fill_Activity$steps <- mapply(filling_values, fill_Activity$steps, fill_Activity$interval)
```

Make a histogram of the total number of steps taken each day and, calculate and report the mean and median total number of steps taken per day.

```{r, echo = TRUE}
totsteps <- tapply(fill_Activity$steps, fill_Activity$date, FUN = sum)
qplot(totsteps, binwidth = 1000, xlab = "Total Number of Steps Taken per Day")
mean(totsteps)
median(totsteps)
```

Do these values differ from the estimates from the first part of the assignment? What is the impact of imputing missing data on the estimates of the total daily number of steps?

The values of the mean and median are higher than the ones calculated in the first part of the assignment. The missing values NA that used to make the total number of steps 0, will be removed since NA's were imputed with the average daily activity pattern.

## Fourth question: Are there differences in activity patterns between weekdays and weekends?

For this part lets use the weekdays() function to distinguish weekdays from weekends in the dataset; the filled-in missing values dataset will be used for this part: `fill_Activity´.

```{r, echo = TRUE}
day.end <- function(date) {
  day <- weekdays(date)
  if (day %in% c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday"))
    return("weekday")
  else if (day %in% c("Saturday", "Sunday"))
    return("weekend")
  else
    stop("invalid date")
}

fill_Activity$date <- as.Date(fill_Activity$date)
fill_Activity$day <- sapply(fill_Activity$date, FUN = day.end)
```

Creating a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).

```{r, echo = TRUE}
Average <- aggregate(steps ~ interval + day, data = fill_Activity, mean)
ggplot(Average, aes(interval, steps)) + geom_line() + facet_grid(day ~ .) + xlab("5-min interval") + ylab("Number of Steps")
```

## THANK YOU!








